# Data Types and Variables

## Data Types
Now, let’s talk about how data are saved in R. R classifies data into four categories: Number, Text, Boolean, Special.

+--------------+-----------------------------------------------------------------------+
| Type         | Description                                                           |
+==============+=======================================================================+
| Number       | Any numbers. They're classified into two categories.                  |
|              |                                                                       |
|              | Integer – Type of number WITHOUT decimals. \                          |
|              | E.g.) 1, 10, 100\                                                     |
|              |                                                                       |
|              | Float – Type of number WITH decimals. \                               |
|              | E.g) 1.24, 2.00 (Not int because of “.00”!)                           |
+--------------+-----------------------------------------------------------------------+
| Text         | Anything you put into “ “, regardless of what’s inside. \             |
|              | E.g) “Hi”, “121”                                                      |
+--------------+-----------------------------------------------------------------------+
| Boolean      | A logic data. It’s either “True” or “False”                           |
+--------------+-----------------------------------------------------------------------+
| Special      | Values that cannot be specified                                       |
|              |                                                                       |
|              | Null – Empty value without determined data type (Number/Text/Boolean) |
|              |                                                                       |
|              | NA – Missing value. Known data type but nothing inside.               |
|              |                                                                       |
|              | NaN – Unable to mathematically define. E.g.) Saving “Hi” as int/float |
|              |                                                                       |
|              | \                                                                     |
|              | inf – Yes, infinite number. They’re under special characters!         |
+--------------+-----------------------------------------------------------------------+


## Variables
Now we know that there are multiple types of data. But how do we tell R to remember them? That’s where variables come in! Variables are names that store data values. 

You can declare a variable by using the assignment operator '<-' or '='. 
```{r}
Test1 <- 3
Test2 = 4
```

Once you declare the variable, you'll see them in the Environment tab. Also, you'll be able to call them by simply typing their names.

```{r}
Test1
Test2
```

If it’s not working, it’s probably due to variable names. There’re specific rules in naming
-	First character must be dot(.) or letter. It cannot be number or special character. Second character and later can be anything
-	You cannot use other function's name as variable name (sum, mean, etc. We'll cover functions as we go on)
-	Variable names are case-sensitive. “test” and “Test” are different variables.
-	Avoid using spaces in variable names. Use dot(.) or underscore(_) instead. E.g.) my.variable, my_variable