# Data Types and Variables

## Data Types

Now, let’s talk about how data are saved in R. R classifies data into four categories: Number, Text, Boolean, Special.

+--------------+-----------------------------------------------------------------------+
| Type         | Description                                                           |
+==============+=======================================================================+
| Number       | Any numbers. They're classified into two categories.                  |
|              |                                                                       |
|              | Integer – Type of number WITHOUT decimals.\                           |
|              | E.g.) 1, 10, 100\                                                     |
|              |                                                                       |
|              | Float – Type of number WITH decimals.\                                |
|              | E.g) 1.24, 2.00 (Not int because of “.00”!)                           |
+--------------+-----------------------------------------------------------------------+
| Text         | Anything you put into “ “, regardless of what’s inside.\              |
|              | E.g) “Hi”, “121”                                                      |
+--------------+-----------------------------------------------------------------------+
| Boolean      | A logic data. It’s either “True” or “False”                           |
+--------------+-----------------------------------------------------------------------+
| Special      | Values that cannot be specified                                       |
|              |                                                                       |
|              | Null – Empty value without determined data type (Number/Text/Boolean) |
|              |                                                                       |
|              | NA – Missing value. Known data type but nothing inside.               |
|              |                                                                       |
|              | NaN – Unable to mathematically define. E.g.) Saving “Hi” as int/float |
|              |                                                                       |
|              | \                                                                     |
|              | inf – Yes, infinite number. They’re under special characters!         |
+--------------+-----------------------------------------------------------------------+

## Variables

Now we know that there are multiple types of data. But how do we tell R to remember them? That’s where variables come in! Variables are names that store data values.

You can declare a variable by using the assignment operator '\<-' or '='.

```{r}
Test1 <- 3
Test2 = 4
```

Once you declare the variable, you'll see them in the Environment tab. Also, you'll be able to call them by simply typing their names.

```{r}
Test1
Test2
```

If it’s not working, it’s probably due to variable names. There’re specific rules in naming - First character must be dot(.) or letter. It cannot be number or special character. Second character and later can be anything - You cannot use other function's name as variable name (sum, mean, etc. We'll cover functions as we go on) - Variable names are case-sensitive. “test” and “Test” are different variables. - Avoid using spaces in variable names. Use dot(.) or underscore(\_) instead. E.g.) my.variable, my_variable

## Simple Calculations

Here, we'll review basic operators in program language.

| Type           | Operation |
|----------------|-----------|
| Addition       | \+        |
| Subtraction    | \-        |
| Multiplication | \*        |
| Remainder      | \%        |
| Exponents      | /         |

Try some on the script you made! Also, you can use variables in calculations.

```{r}
Test1 + Test2
```

There are multiple functions that will make calculations easy

+----------------------+--------------------------------------------------------------+
| **Operation**        | **Description**                                              |
+----------------------+--------------------------------------------------------------+
| ceiling(x)           | Round x up as “int”\                                         |
|                      | E.g.) 3.4 (float) becomes 3 (int)                            |
+----------------------+--------------------------------------------------------------+
| floor(x)             | Round x down as “int”                                        |
+----------------------+--------------------------------------------------------------+
| trunc(x)             | Drop decimals as “float”\                                    |
|                      | E.g.) 3.4 (float) becomes 3.0 (float)                        |
+----------------------+--------------------------------------------------------------+
| round(x, digits = n) | Round x as “float” at nth decimal place.\                    |
|                      | default (If you omit "digits =n") is to round at ones place. |
+----------------------+--------------------------------------------------------------+
| sqrt(x)              | $\sqrt{x}$                                                   |
+----------------------+--------------------------------------------------------------+
| exp(x)               | $e^x$                                                        |
+----------------------+--------------------------------------------------------------+
| log(x, base = n)     | $\log_{n}(x)$. Default is base with n = e                    |
+----------------------+--------------------------------------------------------------+
| sin(x), cos(x), etc. | Trig functions!                                              |
+----------------------+--------------------------------------------------------------+
| Factorial(n)         | n! = 1 x 2 x 3 x ... x n                                     |
+----------------------+--------------------------------------------------------------+
| Others               | There are many other function!                               |
+----------------------+--------------------------------------------------------------+

## Logical Calculations

Sometimes you need more than simple calculations, such as "which number is bigger?" Return values of logical calculations are Boolean (True/False).

| Type       | Operation        |
|------------|------------------|
| Comparison | \<, \>, \<=, \>= |
| Equal      | ==               |
| Not Equal  | !=               |
| Or         | \|               |
| And        | &                |
